---
title: "Introduction to biparty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to biparty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
library(biparty)
library(dplyr)
```

## What `biparty` provides

The `biparty` package provides **Congressional Bipartisanship Scores (CBS)**:
member-level measures of cross-party collaboration in the U.S. Congress,
along with convenience functions for lookup, ranking, issue-area filtering,
and visualization. The scores are introduced and described in Dobson and
Lollis (2026), with support from the Portman Center for Policy Solutions at
the University of Cincinnati.

For every member in every Congress and chamber, two complementary scores
are reported:

- **Attract** — the share of out-party original cosponsors on bills the
  member *sponsored*. A high attract score indicates the member's own
  legislation draws cross-party support.
- **Offer** — the share of the member's own original cosponsorships that
  went to out-party-sponsored bills. A high offer score indicates the member
  actively lends support to the other party's legislation.

Both scores are computed overall and within 34 CRS policy areas. In addition
to the raw (unweighted) measures, the authors provide a weighted version that
adjusts for imprecise measurement associated with legislators with low bill
activity, pulling low-volume observations toward a Congress × chamber (or
Congress × chamber × issue) prior mean. Scores range from 0 to 1, with
higher scores indicating more cross-party activity.

## The two datasets

The package provides two datasets. `aggregate.cbs` contains overall attract
and offer scores for each member-Congress-chamber observation:

```{r}
data(aggregate.cbs)
head(aggregate.cbs)
```

`issue.area.cbs` contains the same scores computed separately within each
of the 34 CRS policy areas:

```{r}
data(issue.area.cbs)
# Preview a few identifier and issue-area columns
issue.area.cbs %>%
  select(congress, chamber, name, party,
         attract_health_weighted, offer_health_weighted) %>%
  head()
```

## Looking up a member

Use `get_member_scores()` to retrieve a member's full score history.
Members can be looked up by Bioguide ID or name substring:

```{r}
# Charles Grassley (R-IA) by Bioguide ID — full career
get_member_scores("G000386")

# Steny Hoyer (D-MD) by name substring
get_member_scores("hoyer", congress = 117)
```

Pass `include_issues = TRUE` to attach issue-area scores:

```{r, eval = FALSE}
grassley_full <- get_member_scores("G000386", congress = 117,
                                   include_issues = TRUE)
```

Use `get_member_trend()` to return a clean time series ready for analysis
or plotting:

```{r}
get_member_trend("G000386", score_type = "attract")
```

## Comparing members

`compare_members()` returns scores for two or more members side by side:

```{r}
compare_members(c("G000386", "H000874"), congress = 115:117)
```

## Ranking members

`rank_members()` returns a leaderboard sorted by attract or offer score
within a given Congress and chamber:

```{r}
# Top 10 senators by attract score in the 100th Congress
rank_members(100, "senate", score_type = "attract")
```

Restrict to one party or rank by an issue-area score:

```{r}
# Top 5 House Democrats by offer score, 117th Congress
rank_members(117, "house", score_type = "offer", party = "D", n = 5)

# Top 10 senators by health attract score, 117th Congress
rank_members(117, "senate", score_type = "attract", issue = "health")
```

## Issue-area analysis

`list_issues()` returns the full crosswalk of short labels to CRS policy
area names:

```{r}
list_issues() %>% head(10)
```

`get_issue_scores()` returns every member's scores for a single policy area:

```{r}
health <- get_issue_scores("health", congress = 117)
head(health)
```

Because `get_issue_scores()` returns a tidy data frame, it can be piped
directly into dplyr or ggplot2 workflows:

```{r}
health %>%
  group_by(chamber, party) %>%
  summarise(
    mean_attract = mean(attract, na.rm = TRUE),
    mean_offer   = mean(offer,   na.rm = TRUE),
    n            = n(),
    .groups      = "drop"
  )
```

Use `get_issue_trend()` to track how a policy area's mean bipartisanship
has changed over time:

```{r}
get_issue_trend("health", chamber = "senate", by_party = TRUE)
```

## Congress-level summaries

`get_congress_summary()` aggregates scores to the Congress × chamber level:

```{r}
# Senate means by party, 115th through 118th Congresses
get_congress_summary(chamber = "senate", congress = 115:118,
                     by_party = TRUE)
```

## Visualization

`plot_bipartisanship()` generates a violin plot of issue-area score
distributions by party and chamber, with an option to highlight a member:

```{r, fig.height = 7}
plot_bipartisanship(100, chambers = "senate", highlight = "G000386")
```

`plot_trend()` produces a time-series line chart of scores across Congresses.
Use `type = "member"` to trace individual careers or `type = "party"` for
party-average trends:

```{r}
# Charles Grassley and Steny Hoyer — full careers
plot_trend(c("G000386", "H000874"), score_type = "attract")
```

```{r}
# Party-average attract trend in the Senate
plot_trend(c("D", "R"), type = "party", chamber = "senate",
           score_type = "attract")
```

`plot_member_profile()` generates a profile card combining a member photo,
career summary statistics, and a bar chart of issue-area scores. The
`magick` and `patchwork` packages are required for the full card layout:

```{r, fig.height = 8}
# Steny Hoyer's profile in the 100th Congress
plot_member_profile("H000874", congress = 100, chamber = "house")
```

## Citation

If you use these scores, please cite the paper:

> Dobson, M.R. and Lollis, J.M. (2026). Congressional bipartisanship scores
> by member and issue area, 1983--2024. Working paper.

and the R package:

> Dobson, M.R. and Lollis, J.M. (2026). *biparty: Congressional
> Bipartisanship Scores by Member and Issue Area*. R package version 0.1.0.
